Revisiting~the~Classics: Online~RL~in~the~Programmable~Data-Plane

### 2021-04-20 paper/paper.tex
More work on design section, starting evaluation.

% Copyright
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\settopmatter{printacmref=false, printccs=false, printfolios=true}

@@

%Conference
\acmConference[SOSR '21]{Symposium of SDx Research 2021}{\numrange{13}{14} September}{Virtual Conference}
%\acmYear{2021}
%\copyrightyear{}

%% {} with no args suppresses printing of the price
\acmPrice{}

@@

\begin{document}
	\title{Revisiting the Classics: Online RL in the Programmable Data-Plane}

	%\titlenote{Produces the permission block, and copyright information}
	%\subtitle{Extended Abstract}

@@

\begin{abstract}
Note: below text is just the UK systems workshop abstract I submitted.	
\end{abstract}

@@

\section{Motivation}\label{sec:motivation}
\fakepara{Reinforcement learning in asynchronous environments}
RL techniques allow an \emph{agent} to learn how to iteratively interact with a system in an optimal fashion.
However, there is some degree of divergence between their theory and implementation.
Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.

@@

\section{Background}

\subsection{Classical Reinforcement Learning}
?? Move into background parent section?

?? RL~\parencite{RL2E}

\subsection{Hmm..}
Should motivation be in here? Unsure what other preliminaries need to be explained.

\section{Design and Implementation}\label{sec:design}
Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{}}.
At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.

\Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.

@@

			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
	\end{tikzpicture}}
	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA)\label{fig:netro-arch}}
\end{figure}

\subsection{System Model}
\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any reinforcement learning agent design} using classical methods.
The meaning 
However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.
\approachshort{} runs on one more cores of a SmartNIC to convert state measurements from the environment into a stream of actions using a stored policy.
These dedicated cores (\emph{micro-engines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.

@@

Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.

?? Dynamic, runtime configuration via control and data planes.

?? Compile-time config of bit-depth, 
?? Allows learning from several live traces simultaneously.
?? Allows execution of indepdnnt traces simultaneously.

\subsection{Action Computation}
Note: talk about single core AND parallel.

?? Mention that you compute action, send action, AND THEN compute update exactly as in that one paper to reduce latency

\begin{itemize}
	\item A state vector is tile coded, action probabilities are produced, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
	\item \emph{Open question: how to configure action destination? This should either be another core (after an optional transformation), a packet sent out to a destination, maybe talk back to the controller.}
\end{itemize}
\begin{itemize}
	\item \emph{Open question: how to safely \& reasonably map state back to incoming rewards?}
	\item \emph{Open question: Can/should this also support on-chip reward signals?}
\end{itemize}

?? Split across MEs.

?? each ME can then subdivide by factor of 8 among contexts.

?? Scales from 1--(n-1) cores.

@@

@@

@@

@@

@@

@@

\subsection{Experimental Setup}

@@

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-work}
	}
	\caption{?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
\end{figure}

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-core}
	}
	\caption{?? Max-size work (129 work-items). ?? Need to dedicate 3 worker to beat independent cores. ?? Sharper perf increase when new physical core added. ?? up to 31 due to master thread.}
\end{figure}

@@

\section{What's implemented 'til now?}
System?
\begin{itemize}
	\item Core-to-core communication.
	\item P4 parser for relevant packets.
	\item Packet copy + handoff from P4 to RL MEs.
	\item Dynamic base configuration.
	\item Dynamic tiling configuration.
	\item Policy installation via block-copy.
	\item Packet generation for above config functions, as well as policy quantisation and splitting.
	\item Tile-coding of quantised states on-chip, and state packet receipt.
	\item Action computation on-chip.
	\item Rudimentary, single-core training.
\end{itemize}

Experiments?
\begin{itemize}
	\item Policy quantisation divergence \& accuracy (via mininet).
	\item Measure DPDK SmartNIC RTT as a lower bound on controller access time.
	\begin{itemize}
		\item Same mechanism applies to other manufacturers (Barefoot), so generalisable.
		\item My results are non-DPDK, but there is a good study~\parencite{DBLP:conf/sigcomm/NeugebauerAZAL018} on exactly this.
	\end{itemize}
	\item Measure basic on-switch times: ring-ring communication, tile-coding cost, action compute cost, learning cost.
\end{itemize}

Known/understood (but not implemented)?
\begin{itemize}
	\item Figure out PIF rule installation.
	\begin{itemize}
		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
	\end{itemize}
\end{itemize}

@@

>>>>

% Copyright
%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
%\setcopyright{none}
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

%\settopmatter{printacmref=false, printccs=false, printfolios=true}
\settopmatter{printfolios=true}

@@

%Conference
\acmConference[SOSR '21]{Symposium of SDx Research 2021}{September \numrange{13}{14}, 2021}{Online}
\acmYear{2021}
%\copyrightyear{}

%% {} with no args suppresses printing of the price
\acmPrice{15.00}

@@

\begin{document}
	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Data-Plane}

%	\titlenote{Produces the permission block, and copyright information}
	%\subtitle{Extended Abstract}

@@

\begin{abstract}
?? In-NIC execution allows clever use of resources to bring action compute latency  from \SI{1}{\milli\second} to \SI{50}{\micro\second} (2 orders of magnitude!) in real-world policies for DDoS detection/prevention control.
\end{abstract}

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10003033.10003068.10003069</concept_id>
	<concept_desc>Networks~Data path algorithms</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10003033.10003099.10003102</concept_id>
	<concept_desc>Networks~Programmable networks</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10003752.10010070.10010071.10010261</concept_id>
	<concept_desc>Theory of computation~Reinforcement learning</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10010147.10010257.10010258.10010261</concept_id>
	<concept_desc>Computing methodologies~Reinforcement learning</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Networks~Data path algorithms}
\ccsdesc[500]{Networks~Programmable networks}
\ccsdesc[500]{Computing methodologies~Reinforcement learning}

%\keywords{ACM proceedings, \LaTeX, text tagging}

@@

\section{Background and Motivation}\label{sec:motivation}
?? RL~\parencite{RL2E}

?? Tile coding requires div, though can be simplified using binary tile widths (relates down to bitshifts). What other ops?

\fakepara{Reinforcement learning in asynchronous environments}
RL techniques allow an \emph{agent} to learn how to iteratively interact with a system in an optimal fashion.
However, there is some degree of divergence between their theory and implementation.
Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.

@@

\section{Design and Implementation}\label{sec:design}
Based on the design principles, problems, and potential solutions outlined throughout \cref{sec:motivation}, we present our design for an in-NIC, task-independent, online reinforcement learning system---\emph{\approachshort{} (\approach)}.
At a high level, \approachshort{} is designed to use the auxiliary compute exposed by general SmartNIC devices to offer low-latency online learning, scaling according to available on-chip resources at build time.

\Cref{fig:netro-arch} outlines our design and implementation on Netronome SmartNIC hardware in pursuit of this goal.
We open-source our firmware design and control programs for the benefit of the community\footnote{\url{https://github.com/FelixMcFelix/pdp-rl-paper}---currently private...}.
We describe how our architectural choices may be adapted and improved upon by more bespoke hardware or FPGA-based deployment.

@@

			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
	\end{tikzpicture}}
	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA, single vs parallel)\label{fig:netro-arch}}
\end{figure}

\subsection{System Model}
\approachshort{} is a general, task-independent framework for in-network, online training and execution of \emph{any reinforcement learning agent design} using classical methods.
\approachshort{} is agnostic to the meaning of state vectors it receives as inputs and the actions it produces, which are employed by other functional units or the dataplane.
However, in-NIC/in-network execution specifically benefits packet-, flow-, and network-level learning, control, and optimisation tasks.

\approachshort{} runs on one more cores of a SmartNIC to convert fixed-point state measurements from the environment into a stream of actions using a stored policy.
These dedicated cores (\emph{microengines} grouped by \emph{island} in NFP terminology) are then responsible for processing requests, computing actions, and updating the underlying policy in real time.
Combined with reward measurements, this policy can then be updated or trained from scratch entirely on the NIC, acting as a fully online RL agent.
An input state vector \emph{always} induces an action, and if desired updates the policy using either an included reward or one retrieved from memory according to a (configurable) key placed alongside the state.
This allows for simultaneous control and learning over independent systems by the same agent (i.e., optimising several flows with their own reward measures).

@@

Runtime reconfiguration and interaction occur via the control and/or dataplane: the ease of use of the P4 pipeline's match and custom protocol parsers, combined with the dedicated input pipe to the NIC's controller (the host machine), allow these to be cleanly separated or combined as needed.
Bitdepth of quantised measurements/preferences, maximum policy sizes, and parallelisation strategy may be configured at compile time.

\subsection{Action and Update Computation}
\approachshort{} applies the insights of \textcite{DBLP:journals/firai/TravnikMSP18} to minimise action latency; an action is computed, sent out into the environment, and only then is the underlying policy updated.
Using one of the below strategies chosen at compile time, a state vector is tile coded, converted into action probabilities, and an action is chosen according to current choice of $\epsilon$ (exploration parameter).
This is then written out to the environment as in \cref{sec:agent-environment-communication}.
If updates are enabled, \approachshort{} then checks an internal hashmap for a previous state-action pair matching the current instruction source, and if found then a policy update is performed.
Updates are computed using \emph{single-step semi-gradient Sarsa}, though modification to support other single-step methods would be trivial.
The new state- or tiles-action pair is then written into storage

\begin{figure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=\linewidth]{figures/single-core-arch-sketch}
		\caption{Single (offline throughput-optimal).\label{fig:single-and-parallel:single}}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=\linewidth]{figures/parallel-arch-sketch}
		\caption{Parallel (online-optimal).\label{fig:single-and-parallel:parallel}}
	\end{subfigure}
	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
\end{figure}

?? Do we need better/snappier names than just Single/Parallel? It's awfully dry.

In the NFP architecture we use, physical cores (MEs) each have 8 hardware threads (contexts): all contexts on an ME share the same program and employ zero-cost context switches on I/O.
Each core then provides 8 worker threads, though in the parallel case one must be dedicated to monitoring work acknowledgements.
Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
\begin{description}
	\item[Single (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially. Computing an action list requires a \emph{read lock} on the policy. If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput. Tile lists are stored for update computation.
	\item[Parallel (\cref{fig:single-and-parallel:parallel})] Threads cooperate on processing state vectors, minimising latency. Workers have a fixed list of work items, while a master thread sends compute/update commands before awaiting worker completion. Work items are disjoint, requiring no policy locks. State vectors are stored for update computation.
\end{description}
Each offers a different point of optimisation; if updates are disabled, then the \emph{single} model can maximise throughput, while the \emph{parallel} model is designed to minimise decision latency and needs no locks for updating the policy (increasing online learning throughput).

We found that, paradoxically, it was more efficient in the parallel case \emph{to do more work} by having each worker recompute its tile subset from a stored state.
It transpired that cacheing this data placed a larger \texttt{memcpy} in the serial section, whose size did not scale at all with bit depth.

@@

?? Reconfig triggers cacheing of per-work item data: tiling locations, ....

Written in rust.

@@

?? Work index correlates with complexity (and can be estimated a-priori).

@@

?? Batching of reads. ?? action value batching  (relate to bitdepth).

@@

\subsection{Limitations}
Known/understood (but not implemented)?
\begin{itemize}
	\item Figure out PIF rule installation.
	\begin{itemize}
		\item Not doable through P4 na\"{i}vely. Handing back rules to controller is expensive: need another action to deliberately use its own hash table etc., then batch back as faster P4 rules.
	\end{itemize}
\end{itemize}

@@

?? Place Dapper~\parencite{DBLP:conf/sosr/GhasemiBR17} on the datapath as the main source of state vectors.

?? Use the TNSM work you reviewed~\parencite{tnms-ddos-victim-ident} as a reward function source.

@@

\subsection{Experimental Setup}
Evaluation of \approachshort{} was performed on server blades hosting a single Netronome Agilio LX 1x100GbE SmartNIC (NFP-6480, \SI{1.2}{\giga\hertz}).
These ran Ubuntu 18.04.5 LTS (4.15.0-140-generic x86\_64) on Intel Xeon Bronze 3204 CPUs (\SI{1.9}{\giga\hertz}) and \SI{32}{\gibi\byte} RAM.
Control programs were built using rustc 1.51.0.
All timing and throughput measurements were repeated over 5000 trials.

Studies based on \textcite{DBLP:journals/tnsm/SimpsonRP20} mirror their experimental setup (Ubuntu 18.04.2 LTS (4.4.3-040403-generic x86\_64), Intel Core i7-6700K (\SI{4.2}{\giga\hertz}), \SI{32}{\gibi\byte} RAM).

@@

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-work}
	}
	\caption{?? More work means More time. ?? Max Parallelisation. ?? Bitdepth affects independent cores more.}
\end{figure}

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-core}
	}
	\caption{?? More cores means less time ?? Max-size work (129 work-items). ?? Need to dedicate 3 worker to beat independent cores. ?? Sharper perf increase when new physical core added. ?? up to 31 due to master thread.}
\end{figure}

@@

@@

\begin{acks}
	This work was supported in part by the \grantsponsor{gs-epsrc}{Engineering and Physical Sciences Research Council}{https://epsrc.ukri.org/} [grant~\grantnum{gs-epsrc}{EP/N509668/1}].
\end{acks}
